---
title: "Line chart"
---

```{r}
#| include: false
library(fontawesome)
```


::: {.callout-tip title="Used geometries"}

**Main geometry:**

- `geom_line`


**Secondary geometries:**

- `geom_point`
- `geom_text_repel`
- `geom_abline`
- `geom_segment`
- `geom_star`
:::






We need to load the following packages:

```{r}
#| message: false
#| warning: false

library(ggrepel)
library(ggtext)
library(ggstar)
library(ggnewscale)

library(here)
library(tidyverse)
```



We load the file `covid_data.csv` in R.

```{r}
#| message: false
#| warning: false

library(readr)
covid_data <- read_csv("data/covid_data.csv", col_types = cols(date = col_date(format = "%m/%d/%Y"))) 

head(covid_data)
```


Let's have a look at the types of variables:

```{r glimpse}
glimpse(covid_data)
```

The data frame contains `r nrow(covid_data)` rows and `r ncol(covid_data)` variables. There are 8 numeric variables, 4 variables of character type, and 2 variables with dates (one of `date` type and the other of `dttm` type):

-   **iso3c:** ISO3c country code as defined by ISO 3166-1 alpha-3
-   **country:** Country name
-   **date:** Calendar date
-   **confirmed:** Confirmed Covid-19 cases as reported by JHU CSSE (accumulated)
-   **deaths:** Covid-19-related deaths as reported by JHU CSSE (accumulated)
-   **recovered:** Covid-19 recoveries as reported by JHU CSSE (accumulated)
-   **total_tests:** Accumulated test counts as reported by Our World in Data
-   **region:** Country region as classified by the World Bank (time-stable)
-   **income:** Country income group as classified by the World Bank (time-stable)
-   **population:** Country population as reported by the World Bank (original identifier 'SP.POP.TOTL', time-stable)
-   **pop_density:** Country population density as reported by the World Bank (original identifier 'EN.POP.DNST', time-stable)
-   **life_expectancy** Average life expectancy at birth of country citizens in years as reported by the World Bank (original identifier 'SP.DYN.LE00.IN', time-stable)
-   **gdp_capita:** Country gross domestic product per capita, measured in 2010 US-\$ as reported by the World Bank (original identifier 'NY.GDP.PCAP.KD', time-stable)
-   **timestamp:** Date and time where data has been collected from authoritative sources



The data cover a period from `r min(covid_data$date)` to `r max(covid_data$date)`.


```{r}
focus_cn <- c("AUS", "BEL", "BRA","CAN","CHN", "FRA", "DEU","GBR",
              "IND", "IRN", "IRL","JPN","ITA", "NLD", "KOR",
               "ESP", "USA")

covid_deaths <- covid_data %>%
  select(date, country, iso3c, deaths) %>%
  group_by(iso3c) %>%
  arrange(date) %>%
  filter(deaths > 9, date < "2020-04-09") %>%
  mutate(days_elapsed = date - min(date),
         end_label = ifelse(date == max(date), country, NA),
         end_label = case_when(iso3c %in% focus_cn ~ end_label,
                      TRUE ~ NA_character_),
         cgroup = case_when(iso3c %in% focus_cn ~ country,
                   TRUE ~ "OTHER")) %>% 
  filter(days_elapsed < 45) |> 
  ungroup()
```



We also set particular colors for the selected countries:

```{r}
# Colors

cgroup_cols <- c("gray70", "gray70", "gray70", "gray70", 
                 "#F69E2F", "#268EC1","gray70","gray70",
                 "gray60", "gray70", "gray30","#7CD6D7",
                 "gray70","#7CD6D7","#298DC3", "#B82E5D", 
                 "#EC5B8F")


# stars
df_stars <- data.frame(
  x_star = c(6.8, 9.7, 10.2, 15, 7.5, 6.5, 2.2, 0.25),
  y_star = c(200, 145, 385, 800, 120, 55, 30, 10),
  group_star = c("a", "a", "b", "c", "d", "d", "e", "d")
)

cgroup_cols2 <- c("#268EC1", "#B82E5D", "gray30", "gray60", "#F69E2F")


# line annotations
df_lines <- data.frame(
  x = c(10, 22, 35, 35),
  y = c(10^4, 2.5*10^4, 3.3*10^4, 350),
  labels = c("DEATHS DOUBLE \n EVERY DAY", "...EVERY \n 2 DAYS",
            "...EVERY \n 3 DAYS", "...EVERY WEEK")
)


```




```{r}
#| message: false
#| warning: false
#| label: fig-final2
#| fig-cap: Final chart
#| fig-width: 12
#| fig-height: 8
#| fig-column: page-right

death_log_curves <- covid_deaths %>% filter(cgroup != "OTHER") %>% 
  ggplot(mapping = aes(x = days_elapsed, y = deaths,
                       color = cgroup, label = end_label,
                       group = iso3c)) +
  geom_line(size = 1.2) +
  geom_point(data = covid_deaths |> 
               dplyr::group_by(cgroup) |> filter(days_elapsed==max(days_elapsed),
                                                 cgroup != "OTHER"),
             aes(x = days_elapsed, y = deaths, fill = cgroup), shape = 21, 
             size = 3.3, stroke = 1.0, color = "grey30") +
  geom_text_repel(nudge_x = 0.2,
                  nudge_y = 0.1, size = 4,
                  segment.color = NA) +
  scale_color_manual(values = cgroup_cols) +
  scale_fill_manual(values = cgroup_cols) +  
  geom_abline(intercept = log10(10), slope = log10(1.42), linetype = "dashed", color = "gray", linewidth = 0.8) +
  geom_abline(intercept = log10(10), slope = log10(1.26), linetype = "dashed", color = "gray", linewidth = 0.8) +
  geom_abline(intercept = log10(10), slope = log10(1.105), linetype = "dashed", color = "gray", linewidth = 0.8) +
  geom_segment(aes(x = 0, y = 10,
                   xend = 10, yend = 10000), linetype = "dashed", color = "gray", linewidth = 0.8) + 
  new_scale_fill() +
  geom_star(data = df_stars, aes(x = x_star, y = y_star, fill = group_star), starstroke = 0.9, size = 3.5, inherit.aes = F) +
  scale_fill_manual(values = cgroup_cols2) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_log10(
    label = scales::comma, sec.axis = dup_axis(), limits = c(10, 33000),
    breaks = c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000, 20000)) +
  labs(x = "Number of days Since 10th death →",
       title = "Coronavirus deaths in Italy, Spain, the UK and US are increasing more rapidly than they did in China",
       subtitle = "Cumulative number of deaths, by number of days since 10th deaths \n \ \ \ \ Stars represent national lockdowns *",
       caption = "Data: https://coronavirus.jhu.edu/map.html") +
  theme(panel.background = element_rect(fill = "#FFF1E6"),
        plot.background = element_rect(fill = "#FFF1E6"),
        panel.grid.major = element_line(color = "#EEE0D5"),
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 16, hjust = -0.95),
        plot.subtitle = element_text(size = 13, hjust = -0.11, color = "grey30"),
        axis.title.y = element_blank(),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.position = "none"
        ) +
  annotate("text", x = df_lines$x, y = df_lines$y, label = df_lines$labels, color = "grey50") +
  annotate("text", x = 43, y = 2.5*1000, label = "China", color = "#F69E2F")

death_log_curves

```



